### 2014-06-18 doc/pseudocode.tex
Added pseudocode (from Eduard).

\documentclass[10pt,onecolumn]{article}

\usepackage[cmex10]{amsmath}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algpseudocode}

\begin{document}

\begin{algorithm}
  \caption{Starcode algorithm}
  \label{alg:starcode}
  \begin{algorithmic}[1]
    \State Define: $\tau$
    \State \textbf{Initialize variables:} $seed$, $start = 0$,
    $height$, $seq$, $trie$
    \State \textbf{Initialize containers:} $hits$, $pebbles$
    \State \textsc{read} sequence file
    \State $height \gets$ \textsc{determine} maximum sequence length
    \State \textsc{pad} sequences up to $height$
    \State \textsc{sort} sequences alphabetically
    \State $trie \gets$ \textsc{create} an empty trie of height $height$
    \State \textsc{insert} root node of $trie$ in $pebbles$ at depth 0
    \ForAll{sequences}
    \State $seq \gets$ \textsc{get} next sequence
    \State $seed \gets$ \textsc{length} of shared prefix
    between current and next $seq$
    \State \textsc{clear} hits
    \State \textsc{clear} $pebbles$ at depth $>start$
    \ForAll{$pebbles$ at depth $start$}
    \State $node \gets$ \textsc{get} next node from $pebbles$
    \State \textbf{call} \textsc{poucet}($node$, $seed$, $hits$)
    \EndFor
    \State \textsc{insert} $seq$ path in $trie$
    \State \textsc{process} $hits$ and \textsc{link} matches to $seq$
    \State $start \gets seed$
    \EndFor
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Poucet search algorithm}
  \label{alg:poucet}
  \begin{algorithmic}[1]
    \Procedure{poucet}{$node$, $seed$, $hits$}:
    \State \textsc{compute} $node$-specific column values following NW algorithm
    \ForAll{$child$ nodes in $node$}
    \State \textsc{compute} $child$-specific row values following NW algorithm
    \State \textsc{compute} center value using row and column
    values
    \If{center value $>\tau$} \Comment{Mismatches exceeded.}
    \State \textbf{continue} with next $child$
    \EndIf
    \If{$node$ depth = $height$} \Comment{Hit found.}
    \State \textsc{save} $node$ sequence in $hits$
    \State \textbf{continue} with next $child$
    \EndIf
    \If{$node$ depth $\leq seed$}
    \State \textsc{save} $node$ in pebbles at current depth
    \EndIf
    \State \textbf{call} poucet($child$, $seed$, $hits$)
    \EndFor
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

\end{document}