### 2014-06-30 doc/pseudocode.tex
Updated pseudocode and Lookup table section.

\documentclass[10pt,onecolumn]{article}

\usepackage[cmex10]{amsmath}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algpseudocode}

\begin{document}

\section{lookup search}
The lookup table search algorithm is based on the fact that if a
sequence is divided in $\tau+1$ non-overlapping subsequences and
mutated with $\tau$ mismatches, at least one subsequence must remain
unchanged for any mismatch combination. In other words,
the distance between two sequences is lower bounded by the number of
different subsequences between them. One can take advantage of this
property to build a table containing all the $k$-mers that have been
previously inserted in the trie. When a new sequence is to be queried,
it is first divided in $\tau+1$ subsequences and all them are queried
against the lookup table. If none of the subsequences are found, the
minimum distance between the query sequence and all the sequences in
the trie is at least $\tau+1$ and the trie-search can therefore be
omitted. This technique is especially useful for long input sequences,
in which the length of the subsequences ($k$) is big and the
probability of finding a coinciding succession of $k$ nucleotides is
only significant when the two sequences do actually belong to the same
cluster. For the rest of the cases, the negative match is detected
only at the computational cost of several table lookups compared to the
enormous cost of measuring the Levenshtein distance between many long
sequences. In our implementation, the length $k$ of the subsequences
is computed in a way that $\tau$ nucleotides are left out. Hence, the
original sequence is divided into $\tau+2$ subsequences, the last of
them containing only $\tau$ nucleotides that will be later used to
quickly detect up to $\tau$ insertions or deletions. The prospective
length of the subsequences is
\begin{equation}
  k = \frac{L}{\tau+1}
\end{equation}
where $L$ is the total length of the sequence. The number of remaining
nucleotides to fill the last subsequence is found as
\begin{equation}
  R = \tau - \big[ L\mod{(\tau+1)} \big]
\end{equation}
and finally the length of each subsequence is computed as
\begin{equation}
  k_i =
  \left\{
  \begin{array}{cl}
    k-1 \qquad & i = 1\ldots R \\
    k \qquad & i = R+1\ldots \tau+1
    \end{array}
    \right.
\end{equation}
The subsequences are inserted in the lookup table after splitting
the original sequence in chunks of length $k_i$. Note that, since the
coincidence with the original sequence must be exact both in position
and content, the rate of false positives returned by the lookup method
is greatly reduced if an independent table is used for each
subsequence. In order to allow for insertions and deletions as candidate
mismatches, the lookup search must also query for shifted versions of
the $k$-mers. Assuming the worst case scenario in which only one of
the subsequences is found and the mismatches are evenly distributed in
the query sequence, the maximum number of cummulative
insertions/deletions at the $i$-th subsequence is exactly $i-1$ ($i =
1\ldots \tau+1$). Hence, the $i$-th subsequence has to be shifted and
queried against the lookup table $2(i-1)+1$ times to cover all the
range from $i-1$ insertions to $i-1$ deletions. The ($\tau+1$)-th
subsequence will therefore be queried $2\tau+1$ times, corresponding
to shifts from $\tau$ nucleotides on the right to $\tau$ nucleotides
on the left, the latter requiring the $\tau$ extra nucleotides left
during the computation of the subsequence lengths.

\begin{algorithm}
  \caption{Starcode algorithm}
  \label{alg:starcode}
  \begin{algorithmic}[1]
    \State \textbf{Define:} $\tau$
    \State \textbf{Variables:} $seed$, $start = 0$,
    $height$, $seq$, $trie$, $lastseq$, $k$
    \State \textbf{Containers:} $hits$, $pebbles$
    \State \textsc{read} sequence file
    \State $height \gets$ \textsc{determine} maximum sequence length
    \State \textsc{pad} sequences up to $height$
    \State \textsc{sort} sequences alphabetically
    \State $k \gets$ \textsc{compute} lookup sequence length
    \Comment{Eq. 3}
    \State $trie \gets$ \textsc{create} an empty trie of height $height$
    \State \textsc{insert} root node of $trie$ in $pebbles$ at depth 0
    \ForAll{sequences}
    \State $seq \gets$ \textsc{get} next sequence
    \If{at least one $k$-mer of $seq$ is in the lookup table}
    \State $seed \gets$ \textsc{length} of shared prefix
    between current and next sequence
    \State $start \gets$ \textsc{length} of shared prefix
    between $seq$ and $lastseq$
    \State \textsc{clear} hits
    \State \textsc{clear} $pebbles$ at depth $>start$
    \ForAll{$pebbles$ at depth $start$}
    \State $node \gets$ \textsc{get} next node from $pebbles$
    \State \textbf{call} \textsc{poucet}($seq$, $node$, $seed$, $hits$)
    \EndFor
    \State \textsc{process} $hits$ and \textsc{link} matches to $seq$
    \State $lastseq \gets seq$
    \EndIf
    \State \textsc{insert} $seq$ path in $trie$
    \State \textsc{insert} $seq$ $k$-mers into the lookup table
    \EndFor
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Poucet search algorithm}
  \label{alg:poucet}
  \begin{algorithmic}[1]
    \Procedure{poucet}{$query$, $node$, $seed$, $hits$}:
    \State \textsc{compute} $node$-specific column values following NW
    algorithm \Comment{Fig.1}
    \ForAll{$child$ nodes in $node$}
    \State \textsc{compute} $child$-specific row values following NW
    algorithm \Comment{Fig.1}
    \State \textsc{compute} center value using row and column values \Comment{Fig.1}
    \If{center value $>\tau$} \Comment{Mismatches exceeded.}
    \State \textbf{continue} with next $child$
    \EndIf
    \If{$node$ depth = $height$} \Comment{Hit found.}
    \State \textsc{save} $node$ sequence in $hits$
    \State \textbf{continue} with next $child$
    \EndIf
    \If{$node$ depth $\leq seed$}
    \State \textsc{save} $node$ in pebbles at current depth
    \EndIf
    \State \textbf{call} poucet($child$, $seed$, $hits$)
    \EndFor
    \EndProcedure
  \end{algorithmic}
\end{algorithm}

\end{document}