### 2015-11-17 doc/tutorial.tex

\documentclass[12pt]{article}
\usepackage{url}
\title{Starcode tutorial}
\author{Guillaume Filion}

\begin{document}
\maketitle

\section{Build instructions}

Starcode runs on Mac and Linux as a command line application.
You can download the source code of Starcode from Github with
the following command on a standard terminal.

\begin{verbatim}
git clone git@github.com:gui11aume/starcode
\end{verbatim}

Note that this requires that you already have a Github account
and that the computer you are working on has an SSH  key registered
on Github. If this is not the case, follow the instructions from
\url{https://help.github.com/articles/generating-ssh-keys/}.

This should download a directory named \texttt{starcode}. To build
Starcode, execute the following.

\begin{verbatim}
cd starcode
make
\end{verbatim}

This should succeed on most Linux systems because \texttt{make} is
available by default. If this is not the case, you can obtain it by
typing \texttt{sudo apt-get install make} on Ubuntu. On Mac, you need
to install XCode, which may take some time. First, you will need an
Apple ID, then you will need to download it from the developer
website of Apple \url{https://developer.apple.com/xcode/downloads/}.
Then, you may need to follow the instructions shown on the following
link to install the command line version of \texttt{make}
\url{http://stackoverflow.com/q/10265742/1248687}.

Calling \texttt{make} should create an executable called
\texttt{starcode}. To check that the building is successful, execute
the following command in the same directory.

\begin{verbatim}
./starcode test/test_file.txt
\end{verbatim}

If you obtain the output shown below, then everything went fine
and you are done with the build. If not, then something went
wrong. In this case, you can explain how to reproduce the problem
on \url{https://github.com/gui11aume/starcode/issues}.

\begin{verbatim}
running starcode with 1 thread
reading input files
raw format detected
sorting
setting dist to 2
progress: 100.00%
AGGGCTTACAAGTATAGGCC  7
CCTCATTATTTGTCGCAATG  7
GGGAGCCCACAGTAAGCGAA  7
TAGCCTGGTGCGACTGTCAT  7
TGCGCCAAGTACGATTTCCG  7
\end{verbatim}

\section{Starcode basics}

Starcode is a sequence clustering algorithm. It takes a bunch of
sequences, finds which are similar and pools them together. Starcode
considers that two sequences are similar if their edit distance
(also known as the Levenshtein distance) is low, \textit{i.e.}
if they almost identical, except for a few mismatches, insertions
and deletions.

The previous example is a standard run of Starcode on sequence
data. By default, Starcode is verbose and gives some information on
how the run is performed before it outputs the clusters. 
In this case, we learn that Starcode used a single CPU, that
the input file \texttt{test/test\char`_file.txt} is in raw format
(\textit{i.e.} plain text), and that the threshold distance for
similarity search was set to 2. The output propoer starts after
the line \texttt{progress: 100.00\%}. It consists of five lines
(one per cluster) where the sequence is the centroid of the cluster
(the irepresentative sequence) and the number is its size (the
number of sequences in the cluster).

The input file \texttt{test/test\char`_file.txt} contains 35 sequences
that Starcode arranged in 5 clusters of 7 sequences each.

We can make Starcode quiet (non verbose) with the \texttt{-q} option,
and we can set the clustering distance to 0 with \texttt{-d0}.

\begin{verbatim}
./starcode -q -d0 test/test_file.txt
\end{verbatim}

The output is the following.

\begin{verbatim}
AGGGCTTACAAGTATAGGCC  6
CCTCATTATTTGTCGCAATG  6
GGGAGCCCACAGTAAGCGAA  6
TAGCCTGGTGCGACTGTCAT  6
TGCGCCAAGTACGATTTCCG  6
AGGGGTTACAAGTCTAGGCC  1
CCTCATTATTTACCGCAATG  1
GGAAGCCCACAGCAAGCGAA  1
TAACCTGGTGCGACTGTTAT  1
TGCGCCAAGTAAGAATTCCG  1
\end{verbatim}

This time we obtain five clusters of size 6 and five clusters of size 1.
Each centroid of the clusters of size 1 has 2 mismatches with one
of the centroids of the clusters of size 6, which is why they form
separate clusters when the distance is less than 2.

As a side note, this example output illustrates that the clusters
are sorted first by size and then by alphabetical order of the
centroid.

\end{document}